---
title: "Reproducible Research: Peer Assessment 2"
output: 
  html_document:
    keep_md: true
---

# Synopsis
An analysis of 
U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.
Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

Across the United States, which types of events have the greatest economic consequences?

For background of project, please refer to [README.md](https://github.com/linearregression/RepData_PeerAssessment2/blob/master/README.md)

## Download data and unzip locally
```{r download, echo=TRUE}
require(R.utils)
targetfile <- 'StormData.csv'
zipData <- paste(targetfile,'.bz2', sep='')

if(!file.exists(targetfile) & !file.exists(zipData)) {
   print('Downloading StormData.zip ....')   
   download.file(url='https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2', destfile=zipData, method='curl')
}
if(file.exists(zipData) & !file.exists(targetfile)) {
   print('Unzip StormData.zip ....')  
   bunzip2(filename=zipData, destname=targetfile, overwrite=TRUE)
}
```
## Loading and preprocessing the data
Load data from csv
```{r loaddata, echo=TRUE}
require(data.table)
Sys.setlocale("LC_TIME", "English")
stormdata <- read.csv2(file=targetfile, header=TRUE, stringsAsFactor=FALSE, na.strings = "NA", colClasses=c("integer", "Date", "integer"))

```
## What is mean total number of steps taken per day?

Frequency distribution of daily total of steps taken. Missing data dropped.

``` {r, echo= TRUE}
stepsFreq <- aggregate(steps~date, data=fitdata, sum, na.rm=TRUE)
colnames(stepsFreq) <- c('date','sum')
hist(stepsFreq$sum, xlab='Daily Steps taken', ylab='Occurence', main='Frequency distribution of daily total number of steps', col='cyan')

```
In additional to the histogram above, group steps by date to vizualize pattern by date. Missing data dropped.
``` {r stepsum, fig.width=4, fig.eight=3, message=FALSE, echo=TRUE, results='hide'}
library(plyr)
library(dplyr)
library(ggplot2)
#fitdata <-na.omit(fitdata)
sumfitdata <- ddply(fitdata,~date, summarise, stepssum=sum(steps, na.rm=TRUE)) 
p <- ggplot(data = sumfitdata, mapping = aes(x=date, y=stepssum)) +
     xlab('Date') +
     ylab('Total Steps') +
     ggtitle('Total Number Steps per Date')
p2 <- p +layer(geom = 'histogram', geom_params = list(fill='green',color = 'steelblue'), stat = 'identity', stat_params = list(binwidth=5))
print(p2)
```
 - Mean number of steps taken per day:
``` {r eval=T, echo=T}
oldmean <- mean(stepsFreq$sum)
oldmean
```   
 - Median number of steps taken per day:

``` {r eval=T, echo=T}
oldmedian <- median(stepsFreq$sum)
oldmedian
```

## What is the average daily activity pattern

Mean number of steps per group by intervals across all days are calculated.
Maximum value of such is also calculated. 

Following is a frequency distribution of Mean number steps per 5 min interval:

``` {r dailyactvity, echo=TRUE}
library(plyr)
library(dplyr)
library(ggplot2)
activitydata <- ddply(fitdata,c('interval'), summarise, 
     mean=mean(steps, na.rm=TRUE)) 
p <- ggplot(data = activitydata, mapping = aes(x=interval, y=mean)) +
     xlab('Interval (min)') +
     ylab('Average Steps') +
     ggtitle('Daily Activity Pattern') 
p <- p+layer(geom="line", stat="identity", geom_params = list(color='red'))
print(p)

```

Interval where Max Mean happens and the corresponding max mean:

``` {r, echo=TRUE}
# Find index on interval axis has the max mean number of steps
index<- which.max(activitydata$mean)
# Interval
activitydata$interval[index]
# Max Mean Steps
activitydata$mean[index]

```


## Imputing missing values

Total number of missing data:

``` {r totalNA, echo=TRUE} 
sum(!complete.cases(fitdata))
```
Strategy to replace missing data. 
Try out replace missing data with mean and median per corresponding interval, then compare changes.
Since mean is senstive to extreme values, try out median.

``` {r meanReplacement, echo = TRUE}
fitdata2<- read.csv(file='activity.csv', header=TRUE, stringsAsFactor=FALSE, na.strings = "NA", colClasses=c("integer", "Date", "integer"))
# Calculate mean and median number of steps per interval

stepmeanbyinterval<- aggregate(steps~interval, data=fitdata2,FUN=mean)
stepsmedianbyinterval <- aggregate(steps~interval, data=fitdata2,FUN=median)
# Replace NA data with mean
fitdata2ByMean <- fitdata2 %>% group_by(interval) %>% mutate(steps= replace(steps, is.na(steps), mean(steps, na.rm=TRUE)))

stepsFreqByMean <- aggregate(fitdata2ByMean$steps, by=list(fitdata2ByMean$date), sum, na.rm=TRUE)
colnames(stepsFreqByMean) <- c('date','sum')
hist(stepsFreqByMean$sum, xlab='Daily Steps taken', ylab='Occurence', main='Replacement of Missing Value as Mean', col='red')

meanbymean <- mean(stepsFreqByMean$sum)
medianbymean <- median(stepsFreqByMean$sum)
#Difference vs data ignoring NA
meanbymean - oldmean
medianbymean - oldmedian

#Try this again with Replacement with Median per that interval
fitdata2ByMedian <- fitdata2 %>% group_by(interval) %>% mutate(steps= replace(steps, is.na(steps), median(steps, na.rm=TRUE)))
stepsFreqByMedian <- aggregate(fitdata2ByMedian$steps, by=list(fitdata2ByMedian$date), sum, na.rm=TRUE)
colnames(stepsFreqByMedian) <- c('date','sum')
hist(stepsFreqByMedian$sum, xlab='Daily Steps taken', ylab='Occurence', main='Replacement of Missing Value as Median', col='yellow')

meanbymedian <-mean(stepsFreqByMedian$sum)
medianbymedian <- median(stepsFreqByMedian$sum)
#Difference vs data ignoring NA
meanbymedian - oldmean
medianbymedian - oldmedian


```
Note that using Replacement by mean strategy does not sway the result that much.

Replacement by median does as you notice most of day is idle except bursts of activity.
So expectation is both lower mean and median.
Replacement by mean assumes 'idle time' during the day is same as daily mean. 


## Are there differences in activity patterns between weekdays and weekends?

```{r weekendWeekday, echo=TRUE}
library(plyr)
library(lattice)

Sys.setlocale("LC_TIME", "English")
fitdata<-na.omit(fitdata)
# Add Weekday Factor variable of .
#Index of rows with weekdays and weekends respectively
fitdays <- weekdays(as.Date(fitdata$date))
weekdays <- (fitdays == 'Saturday') | (fitdays == 'Sunday')
fitdata$isWeekday <- as.factor(ifelse(weekdays, 'weekend', 'weekdays'))
fitdata <- group_by(fitdata, isWeekday)
activitydata <- ddply(fitdata,c('interval','isWeekday'), summarise, 
     mean=mean(steps, na.rm=TRUE))

xyplot(mean ~ interval| isWeekday, data =activitydata, type = "l", layout = c(1, 2),
    xlab = "Interval", ylab = "Mean Number of steps Per Interval", horizontal=FALSE)
```

Both mean and median are different on weekdays vs weekends.

``` {r weekdayVsWeekend, echo=TRUE}

rr <- ddply(fitdata, .(isWeekday, steps, interval), summarise, mean=mean(steps, na.rm=TRUE))
meanwkend <- mean(rr$steps[rr$isWeekday=='weekend'], na.rm=TRUE)
meanwkday <- mean(rr$steps[rr$isWeekday=='weekdays'], na.rm=TRUE)

medianwkend <- median(rr$steps[rr$isWeekday=='weekend'], na.rm=TRUE)
medianwkday <- median(rr$steps[rr$isWeekday=='weekdays'], na.rm=TRUE)

data.frame(day=c('weekday', 'weekend'), 
           mean=c(meanwkday, meanwkend),
           median=c(medianwkday, medianwkend))

```

Publish data

``` {r rpublish, echo=TRUE}

result <- rpubsUpload("My document title", "PA2_template.html")
if (!is.null(result$continueUrl)) 
    browseURL(result$continueUrl) else stop(result$error)

# update the same document with a new title
updateResult <- rpubsUpload("My updated title", "Document.html", result$id)


```

